---
format: 
  html:
    theme: lux
    number-sections: true

title: "In-class Exercise 1"
author: "Tai Yu Ying"
date: "Aug 26 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
---

## Loading the R packages

::: panel-tabset
### The task

For the purpose of this in-class exercise, the following R packages will be used:

-   [tidyverse](https://www.tidyverse.org/)
-   [sf](https://r-spatial.github.io/sf/)
-   [tmap](https://cran.r-project.org/web/packages/tmap/)
-   [ggstatsplot](https://indrajeetpatil.github.io/ggstatsplot/)

Write a code chunk to check if these packages have been installed in R. If yes, load them in R environment.

### The code

```{r}
#| eval: false
pacman::p_load(tidyverse, sf, tmap, ggstatsplot)
```
:::

## Working with Master Plan Planning Sub-zone Data

::: panel-tabset
### The task

1.  Create a sub-folder called data in In-class_Ex01 folder.

2.  If necessary visit data.gov.sg and download Master Plan 2014 Subzone Boundary (Web) from the portal. You are required to download both the ESRI shapefile and kml file.

3.  Write a code chunk to import Master Plan 2014 Subzone Boundary (Web) in shapefile and kml save them in sf simple features data frame.

### The code

*Install the sf Package*

```{r}
options(repos = c(CRAN = "https://cran.rstudio.com/"))
install.packages("sf")
library(sf)
```

This code chunk imports shapefile.

```{r}
mpsz14_shp <- st_read(dsn = "C:/gniyuy/ISSS626_GA/In-class_Ex/In-class_Ex01/Data/",
            layer = "MP14_SUBZONE_WEB_PL")
```

This code chunk imports kml file.

```{r}
#| eval: false
mpsz14_kml <- st_read("C:/gniyuy/ISSS626_GA/In-class_Ex/In-class_Ex01/Data/MasterPlan2014SubzoneBoundaryWebKML.kml")
```
:::

## Working with Master Plan Planning Sub-zone Data

::: panel-tabset
### The task

Write a code chunk to export mpsz14_shp sf data.frame into kml file save the output in data sub-folder. Name the output file MP14_SUBZONE_WEB_PL.

### The code

```{r}
st_write(mpsz14_shp, 
         "C:/gniyuy/ISSS626_GA/In-class_Ex/In-class_Ex01/Data/MasterPlan2014SubzoneBoundaryWebKML.kml",
         delete_dsn = TRUE)
```
:::

## Working with Pre-school Location Data

::: panel-tabset
### The task

1.  If necessary visit data.gov.sg and download Pre-Schools Location from the portal. You are required to download both the kml and geojson files.
2.  Write a code chunk to import Pre-Schools Location in kml geojson save them in sf simple features data frame.

### The code

This code chunk imports kml file.

```{r}
preschool_kml <- st_read("C:/gniyuy/ISSS626_GA/In-class_Ex/In-class_Ex01/Data/PreSchoolsLocation.kml")
```

This code chunk imports geojson file.

```{r}
preschool_geojson <- st_read("C:/gniyuy/ISSS626_GA/In-class_Ex/In-class_Ex01/Data/PreSchoolsLocation.geojson")
```
:::

## Working with Master Plan 2019 Subzone Boundary Data

::: panel-tabset
### The task

1.  Visit data.gov.sg and download Master Plan 2019 Subzone Boundary (No Sea) from the portal. You are required to download both the kml file.
2.  Move MPSZ-2019 shapefile provided for In-class Exercise 1 folder on elearn to data sub-folder of In-class_Ex02.
3.  Write a code chunk to import *Master Plan 2019 Subzone Boundary (No SEA)* kml and MPSZ-2019 into sf simple feature data.frame.

### To import shapefile

```{r}
mpsz19_shp <- st_read(dsn = "C:/gniyuy/ISSS626_GA/In-class_Ex/In-class_Ex01/Data/",
              layer = "MPSZ-2019")
```

### To import kml

```{r}
mpsz19_kml <- st_read("C:/gniyuy/ISSS626_GA/In-class_Ex/In-class_Ex01/Data/MasterPlan2019SubzoneBoundaryNoSeaKML.kml")
```
:::
