---
format: 
  html:
    theme: lux
    number-sections: true

title: "In-class Exercise 9"
author: "Tai Yu Ying"
date: "Sep 28 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
---

## Getting Started

```{r}
pacman::p_load(SpatialAcc, sf, tidyverse, 
               tmap, ggstatsplot)
```

## Count Number of Points within a Distance

### The task

-   To count number of point features (i.e. CHAS clinics) within 1km of another point features (i.e. eldercare centre).

::: callout-note
To complete this section of the in-class exercise, you need to download both the CHAS Clinics and Eldercare Services data sets from data.gov.sg portal. The in-class exercise assumes explicitly the downloaded data sets are saved in rawdata sub-folder of In-class_Ex09 folder. Remember to unzip the file if necessary.
:::

### DIY

-   Download *ELDERCARE* shapefile and *CHAS* kml file from data.gov.sg

-   Using the steps your learned in Hands-on Exercise 1, import *ELDERCARE* shapefile and *CHAS* kml file into R

### The code

*ELDERCARE* is in shapefile format, the code chunk below will be used:

```{r}
eldercare <- st_read(dsn = "data/geospatial",
                     layer = "ELDERCARE") %>%
  st_transform(crs = 3414)
```

The code chunk below is used to import kml file.

```{r}
CHAS <- st_read("data/geospatial/CHAS Clinics.kml") %>%
  st_transform(crs = 3414)
```

### Buffering

Next, `st_buffer()` of sf package is used to create a buffer of 1km around each eldercare features

```{r}
buffer_1km <- st_buffer(eldercare, 
                        dist = 1000)
```

### Visualising

The code chunk below is used to plot the newly created buffers and the CHAS clinics.

```{r}
tmap_mode("view")
tm_shape(buffer_1km) +
  tm_polygons() +
tm_shape(CHAS) +
  tm_dots()
```

### Counting points

Lastly, the code chunk below is used to count the number of CHAS clinics with 1km of each eldercare centre.

```{r}
buffer_1km$pts_count <- lengths(
  st_intersects(buffer_1km, CHAS))
```

## Importing Data

### Geospatial Data

```{r}
mpsz <- st_read(dsn = "Data/geospatial", layer = "MP14_SUBZONE_NO_SEA_PL")

hexagons <- st_read(dsn = "Data/geospatial", layer = "hexagons") 

eldercare <- st_read(dsn = "Data/geospatial", layer = "ELDERCARE") 
```

### OD Matrix

```{r}
ODMatrix <- read_csv("Data/aspatial/OD_Matrix.csv", skip = 0)
```

## Data cleaning and updating attributes

### Supply

```{r}
eldercare <- eldercare %>%
  select(fid, ADDRESSPOS) %>%
  mutate(capacity = 100)
```

### Demand

```{r}
hexagons <- hexagons %>%
  select(fid) %>%
  mutate(demand = 100)
```

### OD Matrix

```{r}
distmat <- ODMatrix %>%
  select(origin_id, destination_id, total_cost) %>%
  spread(destination_id, total_cost)%>%
  select(c(-c('origin_id')))
```

```{r}
distmat_km <- as.matrix(distmat/1000)
```

## Computing Handsen’s Accessibility

### The base code

```{r}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            #d0 = 50,
                            power = 2, 
                            family = "Hansen"))
```

### Tidy the output

```{r}
colnames(acc_Hansen) <- "accHansen"

acc_Hansen <- as_tibble(acc_Hansen)

hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```

### Combine code chunk

```{r}
acc_Hansen <- data.frame(ac(hexagons$demand,
                            eldercare$capacity,
                            distmat_km, 
                            #d0 = 50,
                            power = 0.5, 
                            family = "Hansen"))

colnames(acc_Hansen) <- "accHansen"
acc_Hansen <- as_tibble(acc_Hansen)
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```

### Visualising Accessibility

### The code

```{r}
mapex <- st_bbox(hexagons)

tmap_mode("plot")
tm_shape(hexagon_Hansen,
         bbox = mapex) + 
  tm_fill(col = "accHansen",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: Hansen method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```
